\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-08 paper/draft.tex
Corrections

%% OSDI -- 12 pages limit
%% \documentclass[letterpaper,twocolumn,10pt]{article}
%% \pdfoutput=1 % for arxiv
%% \usepackage{usenix2019_v3}

@@

\tr{\kk{If we talk about the language it would be good to give some
  statistics about how many commands (out of the ones that we have
  classified) can be represented with one, two clauses etc. If we have
  a different solution (like python function that given a command and
  its arguments returns the category) then we could talk about how
  many lines of code it took to categorize all the commands that we
  did.}}

\tr{\kk{Note: Since a categorization language will probably not be
  complete (especially the one showing the input arguments) there
  should be a backup mechanism, when the language is not expressive
  enough, like a python function that categorizes the command, and
  returns its input argument if it is stateless.}}

\tr{\kk{(Maybe) We should have a crisp point about why we have this categorization
  language, and why don't we just allow someone to write a function in
  python for each command, that given the command and its arguments,
  returns whether the command category. Possible arguments include,
  the fact that this language is simpler to use, especially by non
  experts that just run some script but have installed some commands
  that are not ``supported''. Another possible argument is that we
  could be able to reason about the constructs of the categorization
  lagnauge, and that they will be easier to read and
  understand. Another argument is that since these categorizations
  should be shareable among users, it would be bad to execute
  arbitrary python code, so using this language, expressivity is
  limited. All of these arguments are a little bit weak though. Niko,
  what do you think?}}

\heading{Complete Annotation Language}
The complete annotation language contains 6 operators, one of which supports regural expressions.
It was used to annotate 47 commands, totalling 708 lines of JSON---an effort that took about 2--3 hours.
Annotation records are by default conservative so as to not jeopardize correctness, but can be incrementally refined to capture parallelizability when using increasingly complex combinations of flags.
The language is extensible with more operators (as long as the \todo{user} defines an \todo{interpreter} for them);
  it also supports writing arbitrary python code for commands whose properties are difficult to capture---\eg higher order \ttt{xargs}, whose parallelizability class depends on the class of the first-order command that it invokes.
% \str
% \kk{Mention that some commands, such as xargs are handled internally due to their particularities.}

@@

\tr{\kk{I am not sure a general interface is so easy to design. It needs
  more though. It might be beneficial to just talk about sort and wc
  here and how we implemented them and nothing more. Or maybe this
  could then go to the implementation? Or maybe say that one can write
  a python function that given a node of the graph, transforms it into
  many. I am not sure what is best...}}

@@

\heading{Highlights}
\sys accelerates ($1.92-61.1\times$) all scripts ($>50$, with $>200$ commands).
For some of our benchmarks~\sx{ours}, we decouple the performance benefits stemming from \sys's compiler alone from the ones stemming from \sys's runtime primitives:
  the compiler's average speedup seen is $9.3\times$, jumping to $13.2\times$ when runtime optimizations are used.
In a different set of $30$ benchmarks parallelized $16\times$\footnote{
  \nv{Immediate question: why use $64\times$ in one and $16\times$ in another?}
}
 see an average speedup of $6\times$, depending on the class of piecewise commands~\sx{unix50}.
Two complex scripts with 15~\sx{macro1} and 34~\sx{macro2} commands in multiple languages see $2.44\times$ and $12.7\times$ speedup on $16\times$-parallel configuration.
\sys outperforms \ttt{sort --parallel} by a significant margin, and offers small performance benefits over GNU \ttt{parallel}, which returns incorrect results in one configuration~\sx{micro}.
% Averages: [1.955223776839714, 3.478591807121036, 5.721328556478535, 8.604380509595906, 10.666921423438112, 13.246864076708423]
% No Eager Averages: [1.6327834370824543, 2.5481043878162444, 3.8623841529643173, 5.931851620007484, 7.466566647963421, 9.35752032439642]
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
% \sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
% In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

The vast majority of programs require no effort to parallelize other than invoking \sys;
  only 6 commands ($<3\%$) needed a single-record annotation.
% 
In terms of correctness, \sys's results over multi-GB  inputs are identical to the sequential for all benchmarks.
Scripts feature ample opportunities for breaking semantics, which \sys avoids.
%% In one case, \sys avoids transforming a cycle (\ie a fragment whose input and output identifiers are the same).
%in another, it preserves the semantics that another parallelization command breaks.

@@

We evaluate \sys under multiple configurations on popular, common, or classic \unix pipeline patterns~\cite{bentley1985spelling, bentley1986literate, taylor2004wicked}.

@@

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of parallelism (2--64$\times$).
% (Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups of the optimized \sys, \ie \ttt{eager} and \ttt{split} enabled, for \{2, 4, 8, 16, 32, 64\}$\times$-parallelism are \todo{1.97, 3.5, 5.78, 8.83, 10.96, and 13.47}, respectively.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}
% \kk{For the one-liners where split doesn't provide any benefit (or overhead) we don't show the PaSh w/o split line.}
%
% we experiment with a $200\times$ level of parallelism to see potential overheads
Tab.~\ref{tab:eval} shows that \sys's compilation time is negligible. % ($<5s$)
It also contains filter and bigrams-opt (not in Fig.~\ref{fig:microbenchmarks}). The first (filter) is used to evaluate \sys's impact on IO-intensive scripts (\sys achieves a 1.5-2.5$\times$ speedup). The second (bigrams-opt) is an optimized version of bigrams where we grouped 5 commands into one \tpur command thas has a more efficient $map$ and $aggregate$ implementation. \sys achieves up to x30 speedup for this script (significantly higher than what is achieved for the standard bigrams Fig.\ref{fig:microbenchmarks}).
\sys achieves a COST~\cite{mcsherryscalability} of 2 for all benchmarks.
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

\heading{Discussion}
%
As discussed in (\S\ref{eval}, Parallelism) \sys does not have any control over the input script parallelism, and therefore it often achieves the maximum possible speedup, \ie without unnecessary process spawns and context switches, earlier than the purported parallelism---$16-32\times$ for a 64-core system. This is also due to the fact that \sys also adds its runtime primitives (\ttt{split} and \ttt{eager}), further increasing process spawns.
% \kk{eurosys: Can we somehow turn this point around and make it seems good/fine. That here we report this parallelism, but in the real system, it tries to do a best effort guess of parallelism based on the size of the input script.}
% \sys achieves near-linear speedup on low-parallelism configurations.
% This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \emph{aggregation} nodes, and 14 \ttt{relay} nodes.
% In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
% significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
% Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).
% \kk{Mention bigrams-opt result now that it is removed from the plot.}
\todo{No Eager} and \todo{Blocking Eager} depict configurations without \sys's \ttt{eager} optimization and without \ttt{split}. As described in \Cref{optimizer}, these eager implementations perform worse than \sys's optimized relay \ttt{eager} in all cases (except for a negligible difference in shortest-scripts). The benefits of \ttt{split} are also eveident for scripts that have \tpur or \tnpu commands, and for the rest it does not negatively affect performance.

@@

\heading{Results}
The complete pipeline executes in 44m2s for five years (82GB) of data.
\sys with 2/10$\times$ parallelism leads to 1.86/2.44$\times$ speedup, with different phases seeing different benefits:
  1.77/1.99$\times$ (vs. 33m58s) for all the pre-processing and 
  2.30/10.79$\times$ speedup (vs. 10m4s) for computing the max.
The reason why speedup is higher than parallelism is explained at the beginning of \S\ref{eval} (Parallelism).\footnote{
  Can we avoid this by not showing the $2\times$? Why show $2\times$?
}

>>>>

% TODO Add Ignis and Mozart
%% \documentclass[letterpaper,twocolumn,10pt]{article}
%% \pdfoutput=1 % for arxiv
%% \usepackage{usenix2019_v3}

@@

\tr{\kk{If we talk about the language it would be good to give some
  statistics about how many commands (out of the ones that we have
  classified) can be represented with one, two clauses etc. If we have
  a different solution (like Python function that given a command and
  its arguments returns the category) then we could talk about how
  many lines of code it took to categorize all the commands that we
  did.}}

\tr{\kk{Note: Since a categorization language will probably not be
  complete (especially the one showing the input arguments) there
  should be a backup mechanism, when the language is not expressive
  enough, like a Python function that categorizes the command, and
  returns its input argument if it is stateless.}}

\tr{\kk{(Maybe) We should have a crisp point about why we have this categorization
  language, and why don't we just allow someone to write a function in
  Python for each command, that given the command and its arguments,
  returns whether the command category. Possible arguments include,
  the fact that this language is simpler to use, especially by non
  experts that just run some script but have installed some commands
  that are not ``supported''. Another possible argument is that we
  could be able to reason about the constructs of the categorization
  lagnauge, and that they will be easier to read and
  understand. Another argument is that since these categorizations
  should be shareable among users, it would be bad to execute
  arbitrary Python code, so using this language, expressivity is
  limited. All of these arguments are a little bit weak though. Niko,
  what do you think?}}

\heading{Complete Annotation Language}
The complete annotation language contains 6 operators, one of which supports regural expressions.
It was used to annotate 47 commands, totalling 708 lines of JSON---an effort that took about 2--3 hours.
Annotation records are by default conservative so as to not jeopardize correctness, but can be incrementally refined to capture parallelizability when using increasingly complex combinations of flags.
The language is extensible with more operators (as long as the \todo{user} defines an \todo{interpreter} for them);
  it also supports writing arbitrary Python code for commands whose properties are difficult to capture---\eg higher order \ttt{xargs}, whose parallelizability class depends on the class of the first-order command that it invokes.
% \str
% \kk{Mention that some commands, such as xargs are handled internally due to their particularities.}

@@

\tr{\kk{I am not sure a general interface is so easy to design. It needs
  more though. It might be beneficial to just talk about sort and wc
  here and how we implemented them and nothing more. Or maybe this
  could then go to the implementation? Or maybe say that one can write
  a Python function that given a node of the graph, transforms it into
  many. I am not sure what is best...}}

@@

\heading{Highlights}
\sys accelerates ($1.92-61.1\times$) all scripts.\footnote{
  \nv{But this is not true; ``the vast majority''? maybe put 51/54 scripts?}.
}
For some of our benchmarks~\sx{ours}, we decouple the performance benefits stemming from \sys's compiler alone from the ones stemming from \sys's runtime primitives:
  the compiler's average speedup seen is $9.3\times$, jumping to $13.2\times$ when runtime optimizations are used.
In a different set of $30$ benchmarks parallelized $16\times$\footnote{
  \nv{Immediate question: why use $64\times$ in one and $16\times$ in another?}
}
 see an average speedup of $6\times$, depending on the class of piecewise commands~\sx{unix50}.
Two complex scripts with 15~\sx{macro1} and 34~\sx{macro2} commands in multiple languages see $2.44\times$ and $12.7\times$ speedup on $16\times$-parallel configuration.
\sys outperforms \ttt{sort --parallel} by a significant margin, and offers small performance benefits over GNU \ttt{parallel}, which returns incorrect results in one configuration~\sx{micro}.
% Averages: [1.955223776839714, 3.478591807121036, 5.721328556478535, 8.604380509595906, 10.666921423438112, 13.246864076708423]
% No Eager Averages: [1.6327834370824543, 2.5481043878162444, 3.8623841529643173, 5.931851620007484, 7.466566647963421, 9.35752032439642]
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
% \sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
% In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

The vast majority of programs ($>50$, with $>200$ commands) require no effort to parallelize other than invoking \sys;
  only 6 commands ($<3\%$) needed a single-record annotation.
% 
In terms of correctness, \sys's results over multi-GB  inputs are identical to the sequential for all benchmarks.
Scripts feature ample opportunities for breaking semantics, which \sys avoids.
%% In one case, \sys avoids transforming a cycle (\ie a fragment whose input and output identifiers are the same).
%in another, it preserves the semantics that another parallelization command breaks.

@@

We evaluate \sys under multiple configurations on popular, common, or classic \unix pipeline patterns~\cite{bentley1985spelling, bentley1986literate, taylor2004wicked}.
% \nv{Eurosys: goal of experiment: decouple compiler from runtime, and talk about all overheads}

@@

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of parallelism (2--64$\times$).
% (Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups of the optimized \sys, \ie \ttt{eager} and \ttt{split} enabled, for \{2, 4, 8, 16, 32, 64\}$\times$-parallelism are \todo{1.97, 3.5, 5.78, 8.83, 10.96, and 13.47}, respectively.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}
% \kk{For the one-liners where split doesn't provide any benefit (or overhead) we don't show the PaSh w/o split line.}

% we experiment with a $200\times$ level of parallelism to see potential overheads
Tab.~\ref{tab:eval} shows that \sys's compilation time is negligible. % ($<5s$)
It also contains Filter and Bigrams-opt (not in Fig.~\ref{fig:microbenchmarks}).
Filter is used to evaluate \sys's impact on IO-intensive scripts (\sys achieves a 1.5--2.5$\times$ speedup).
Bigrams-opt is an optimized version of Bigrams that groups 5 commands into one \pur command and thus has a more efficient $map$ and $aggregate$ implementation.
\sys achieves up to $30\times$ speedup for this script (significantly higher than what is achieved for the standard Bigrams Fig.\ref{fig:microbenchmarks}).
% \nv{Eurosys: I think this should be part of a "do custom aggregators help?" question, not the main results}.
\sys achieves a COST~\cite{mcsherryscalability} of 2 for all benchmarks.
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

\heading{Discussion}
%
\todo{These results explain a detail before getting the main points across: A is better than B, C is worse except G}
As discussed earlier (\S\ref{eval}, Parallelism), \sys does not have control over the input script's parallelism.
Thus, it often achieves the maximum possible speedup, \ie without unnecessary process spawns and context switches, earlier than the purported parallelism---$16-32\times$ for a 64-core system.
This is also due to the fact that \sys also adds its runtime primitives (\ttt{split} and \ttt{eager}), further increasing process spawns.
% \kk{eurosys: Can we somehow turn this point around and make it seems good/fine. That here we report this parallelism, but in the real system, it tries to do a best effort guess of parallelism based on the size of the input script.}
% \sys achieves near-linear speedup on low-parallelism configurations.
% This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \emph{aggregation} nodes, and 14 \ttt{relay} nodes.
% In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
% significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
% Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).
% \kk{Mention bigrams-opt result now that it is removed from the plot.}
\todo{No Eager} and \todo{Blocking Eager} depict configurations without \sys's \ttt{eager} optimization and without \ttt{split}.
As described in Fig.\ref{optimizer}, these eager implementations perform worse than \sys's optimized \ttt{eager} relay in all cases (except for a negligible difference in Shortest-Scripts).
A \ttt{split} benefits scripts with \tpur or \tnpu commands, but does not negatively affect the rest.

@@

\heading{Results}
The complete pipeline executes in 44m2s for five years (82GB) of data.
\sys with 2/10$\times$ parallelism leads to 1.86/2.44$\times$ speedup, with different phases seeing different benefits:
  1.77/1.99$\times$ (vs. 33m58s) for all the pre-processing and 
  2.30/10.79$\times$ speedup (vs. 10m4s) for computing the max.
The reason why speedup is higher than parallelism is explained at the beginning of \S\ref{eval} (Parallelism).\footnote{
  \nv{Can we avoid this by not showing the $2\times$? Why show $2\times$?} % <--- Eurosys
}
